Borel regularity for ODEs and Thimble Integrals

### 2023-12-05 exp-resurgence.tex
Removed old part from Section 2.2

\begin{remark}
There are many ways to build a frame of solutions for these kinds of ODEs: formally in the $z$-plane Poincar\'e gave an explicit solution \cite{int-irreg}. Formally in the position domain \'Ecalle figured out how to find a frame of formal solutions of the Borel transformed ODE \cite{EcalleIII,loday-Remy2011}. It can indeed be seen as a consequence of ``resurgence'' of the ODE that the formal solutiomns can be analytically continued. Our results of \cite{reg-sing-volterra} show how to build a frame of solutions, analytically in the Borel plane. Here, we show we are picking the same frame: from the properties of the Laplace transform we get an analytic frame asymptotic to Poincar\'e frame $\series{\Psi}_j$. From the uniqueness result of \cite[Theorem~4]{reg-sing-volterra}, we get a frame of analytic solutions equivalent to \'Ecalle's ones. 
\end{remark}

\color{gray}

As described in Section~\ref{borel-reg:explanatory-power}, we consider a linear single level-$1$ ODE $\mathcal{P}\Phi = 0$ given by a differential operator of the form
\begin{equation}%\label{eqn:standard ODE}
\mathcal{P} = P(\partial_z) + \frac{1}{z} Q(\partial_z) + \frac{1}{z^2}\big[ R_0(z^{-1}) + R_1(z^{-1})\,\partial_z + \ldots + R_{d-1}(z^{-1})\,\partial_z^{d-1} \big],
\end{equation}
where $P$ is a monic degree-$d$ polynomial, $Q$ is a degree-$(d-1)$ polynomial, and $R_j(z^{-1})$ are holomorphic in some disk $|z| > A$ around $z = \infty$, for every $j=0,\ldots, d-1$.\footnote{Looking at the existence theorem in \cite[Theorem 4]{reg-sing-volterra}, we could apply this reasoning on the analytic side for more general equations, but this particular case makes it easier to talk about the formal side as well.} Furthermore we assume $P$ has simple zeros $P(-\alpha_j)=0$, $j=1,...,d$, $P'(-\alpha_j)\neq 0$ for $j=1,...,d$ and $Q(-\alpha_j)\neq 0$. Our equation has an irregular singularity at $\infty$ (see \cite[Definition 3.3.2]{diverg-resurg--ii}).

Under these assumptions, the space of formal solutions of $\mathcal{P}\Phi = 0$ has a basis $\series{\Psi}_1,...,\series{\Psi}_d$ of the form~\cite{int-irreg}\cite[Proposition~2.2.7, p.~111]{EcalleIII}
\begin{equation}%\label{formal solution}
\series{\Psi}_j(z)=e^{-\alpha_j z}z^{-\tau_j}\series{F}_j(z)\in e^{-\alpha_j z } z^{-\tau_j}\,\C \llbracket z^{-1} \rrbracket,
\end{equation}
where $\tau_j$ is the non-zero complex number $Q(-\alpha_j)/P'(-\alpha_j)$. Since the roots $-\alpha_1, \ldots, -\alpha_n$ are distinct, this form determines the basis up to scaling, giving the space of formal solutions a distinguished frame. \textcolor{orange}{Add that the formal series is $1$-Gevrey.}

Classically it was proved that the formal solutions of \eqref{formal solution} are Borel-Laplace summable and their Borel-Laplace sums satisfy the original differential equation (see for example \cite{ramis1991series,malgrange92,malgrange1995sommation,diverg-resurg--ii}\textcolor{magenta}{[Braaksma,check]}). Therefore, the distinguished frame of formal solutions $\series{\Psi}_1,...,\series{\Psi}_d$ becomes a distinguished frame of analytic solutions $\hat{\Psi}_1,...,\hat{\Psi}_d$.

However, can we find a distinguished basis for solutions in a purely analytic way? This should be possible because of our regularity assumptions about the coefficients of $\mathcal{P}$. In fact, the Main Asymptotic Existence Theorem (M.A.E.T.) guarantees the existence of an analytic frame of solutions asymptotic to a formal frame (see \cite[Theorem 3.1]{diverg-resurg-iii}), but it is not a constructive method (see \cite[Chapter 14]{balser}). Conversely, we can explicitly build a distinguished basis of analytic solutions: first going to the position domain, we can consider $d$ integral operators of the form 
\begin{equation}
\hat{P}_j:=P(-\zeta)+\partial_{\zeta,\alpha_j}^{-1}\circ Q(-\zeta)+\partial_{\zeta,\alpha_j}^{-2}\circ\sum_{r=0}^{d-1}R_r(\partial_{\zeta,\alpha_j}^{-1})\circ (-\zeta)^r
\end{equation}
labeled by the roots of $P$. This type of operator is the one we studied in \cite{reg-sing-volterra}, hence we know there exists a unique solution $\psi_j$ of $\hat{P}_j\psi_j=0$, such that $\psi_j$ blows-up as $\zeta_j^{\tau_j-1}$ at $\zeta_j=0$ and it has a controlled behavior on an open domain $\Omega_j$ which is star-shaped at $\alpha_j$ and avoids the other critical points (see the definition in Section \ref{setting-ODE}). In addition, $\psi_j$ can be Laplace transformed, giving a analytic function $\Psi_j:=\laplace_{\zeta_j}^{\theta}\psi_j$. 
At this point, we may ask what is the relationship between the two frames of analytic solutions, namely the Borel sum $\hat{\Psi}_j$ of the formal solutions $\series{\Psi}_j$ and the Laplace transform of $\psi_j$.

\begin{remark}
There are many ways to build a frame of solutions for these kinds of ODEs: formally in the $z$-plane Poincar\'e gave an explicit solution \cite{int-irreg}. Formally in the $\zeta$-plane \'Ecalle figured out how to find a frame of formal solutions of the Borel transformed ODE \textcolor{magenta}{[add reference]\cite{EcalleIII,loday-Remy2011}}. It can indeed be seen as a consequence of ``resurgence'' of the ODE. Our results of \cite{reg-sing-volterra} show how to build a frame of solutions, analytically in the Borel plane. Here, we show we are picking the same frame: from the properties of the Laplace transform we get an analytic frame asymptotic to Poincar\'e frame $\series{\Psi}_j$. From the uniqueness result of \cite[Theorem~4]{reg-sing-volterra}, we get a frame of analytic solutions equivalent to \'Ecalle's ones. 
\end{remark}

\begin{remark}
The theory of $1$-summability allows us to draw the ``square'' with corners $\hat{\psi}_j$ and its Laplace transform $\hat{\Psi}_j$ (namely, drawing the red arrow of the diagram). This ensures that the Borel sum of the formal solution $\series{\Psi}_j$ exists and it solves the ODE. However, the approach of Borel regularity is different as it is based on the analytic solution in the position domain. 
\end{remark}
\begin{remark}
    Our approach is different from the one of Malgrange in \cite{malgrange--fourier} where he uses differential equations in the \textcolor{red}{spatial/frequency?} domain. 
\end{remark}

\subsection{Borel regularity for thimble integrals}
%
Thimble integrals have been studied from different perspectives: in physics, they play an important technical role in quantum mechanics, where infinite-dimensional exponential integrals are supposed to give the expectation values of observable quantities \cite{dunne-unsal2,dunne-unsal,Fauvet_Menous_Queva,Tanizaki:2014tua}. In this context, physicists often use Borel summation, resurgence and related techniques to assign values to these integrals going beyond their pertubative expansion. For instance, in complex Chern--Simons theory, one could decompose the path integral in thimble integrals and arguing as in the finite dimesional set-up to study the Witten--Reshetikhin--Turaev invariants of $3$ manifolds \cite{gukov-marino-purtrov-resurgence,Witten}. \textcolor{orange}{Take a look at Witten's paper on knots}

>>>>

\begin{remark}
There are many ways to build a frame of solutions for these kinds of ODEs: formally in the $z$-plane Poincar\'e gave an explicit solution \cite{int-irreg}. Formally in the position domain \'Ecalle figured out how to find a frame of formal solutions of the Borel transformed ODE \cite{EcalleIII,loday-Remy2011}. It can indeed be seen as a consequence of ``resurgence'' of the ODE that the formal solutiomns can be analytically continued. Our results of \cite{reg-sing-volterra} show how to build a frame of solutions, analytically in the Borel plane. Here, we show we are picking the same frame: from the properties of the Laplace transform we get an analytic frame asymptotic to Poincar\'e frame $\series{\Psi}_j$. From the uniqueness result of \cite[Theorem~4]{reg-sing-volterra}, we get a frame of analytic solutions equivalent to \'Ecalle's ones. 
\end{remark} 
%
%
%
\subsection{Borel regularity for thimble integrals}
%
Thimble integrals have been studied from different perspectives: in physics, they play an important technical role in quantum mechanics, where infinite-dimensional exponential integrals are supposed to give the expectation values of observable quantities \cite{dunne-unsal2,dunne-unsal,Fauvet_Menous_Queva,Tanizaki:2014tua}. In this context, physicists often use Borel summation, resurgence and related techniques to assign values to these integrals going beyond their pertubative expansion. For instance, in complex Chern--Simons theory, one could decompose the path integral in thimble integrals and arguing as in the finite dimesional set-up to study the Witten--Reshetikhin--Turaev invariants of $3$ manifolds \cite{gukov-marino-purtrov-resurgence,Witten}. \textcolor{orange}{Take a look at Witten's paper on knots}
